\thetitle

### 2021-05-20 writing/challenges.tex
move Louisiana footnote so that the intro to the benchmark section makes sense

The {\em second challenge\/} is to find a representative collection of
programs with impedance mismatches. The latter must represent mistakes that programmers
accidentally create and that the run-time checks of academic systems catch. In
other words, the experiment calls for a collection of mistakes in
mix-typed programs that is representative of those ``in the wild.''
Unfortunately no such collection exists, and with good reason. The kind of
mistakes needed are typically detected by unit or integration tests; even if it
takes some time to find their sources, these mistakes do not make it into code
repositories with appropriate commit messages.

>>>>

The {\em second challenge\/} is to find a representative collection of programs
with impedance mismatches.\footnote{\citet{cc-oopsla-20} created a collection of
Reticulated Python programs to evaluate their technique of fixing mistakes in
type annotations. Their collection does not come with type-level mistakes in the
code itself.}  The latter must represent mistakes that programmers accidentally
create and that the run-time checks of academic systems catch. In other words,
the experiment calls for a collection of mistakes in mix-typed programs that is
representative of those ``in the wild.''  Unfortunately no such collection
exists, and with good reason. The kind of mistakes needed are typically detected
by unit or integration tests; even if it takes some time to find their sources,
these mistakes do not make it into code repositories with appropriate commit
messages.

### 2021-05-20 writing/mutate-benchmarks.tex
move Louisiana footnote so that the intro to the benchmark section makes sense

\item have a decent number of type-able modules and a variety of module dependency graphs
 (because mixing of typed and untyped code in Typed Racket takes place at the
 module level);

\item take advantage of the variety of typing features of a gradually typed
 language; but

\item fail and contain type-level mistakes. 
\end{enumerate}
No such benchmark suite exists.  Worse, no benchmark suite of
gradually typed programs with type-level mistakes exists at
all.\footnote{\citet{cc-oopsla-20} created a collection of Reticulated
Python programs to evaluate their technique of fixing mistakes in type
annotations. Their collection does not come with type-level mistakes
in the code itself.}

What exists is \citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs
for systematically measuring the implementation's performance. The benchmark
suite consists of fully typed, correct programs, written by a number of
different authors, who have maintained and evolved these programs over several
years. The programs range widely in size, complexity, purpose, origin and in
programming style. The latter is mostly a reference to their reliance on many
Typed Racket features: occurrence typing~\citep{tf-icfp-2010}, types for mutable
and immutable data structures~\citep{hpst-sfp-2010}, types for first-class
classes and objects~\citep{tsdtf-oopsla-2012}, and types for Racket's numeric
tower~\citep{stathff-padl-12}.
Finally, all of the programs are deterministic, so any changes in the programs' behavior
between runs can be solely attributed to the actions of the rational programmer.

>>>>

\item be fully typed so that the choice of types is fixed;

\item take advantage of the variety of typing features of a gradually
typed; and

\item have a decent number of type-able modules and a variety of module dependency graphs
 because mixing of typed and untyped code in Typed Racket takes place at the
 module level.

\end{enumerate}

\citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs for
systematically measuring the implementation's performance satisfies these
criteria. The benchmark suite consists of fully typed, correct programs, written
by a number of different authors, who have maintained and evolved these programs
over several years. The programs range widely in size, complexity, purpose,
origin and in programming style. The latter is mostly a reference to their
reliance on many Typed Racket features: occurrence typing~\citep{tf-icfp-2010},
types for mutable and immutable data structures~\citep{hpst-sfp-2010}, types for
first-class classes and objects~\citep{tsdtf-oopsla-2012}, and types for
Racket's numeric tower~\citep{stathff-padl-12}.  Finally, all of the programs
are deterministic, so any changes in the programs' behavior between runs can be
solely attributed to the actions of the rational programmer.

### 2021-05-20 writing/mutate.tex
move Louisiana footnote so that the intro to the benchmark section makes sense

Putting the rational programmer to work means generating many mutants and turning
those into debugging scenarios. The process must start with a suitable
collection of representative programs (see sec.~\ref{sub:mutate-benchmarks}).
Since existing mutators do not generate useful mutants, the first step is to
develop new mutators (see sec.~\ref{sub:mutate-mutators}) and to validate their
suitability on the benchmarks (see sec.~\ref{sub:mutate-interesting}).

% As section~\ref{sec:challenges} explains, the first challenge is to
% generate a representative mistakes in representative software. This
% section explains the specific choices made for the first application
% of the method: the source and nature of the chosen benchmarks
% (sec.~\ref{sub:mutate-benchmarks}), the newly developed mutators
% (sec.~\ref{sub:mutate-mutators}), and why they are suitable for this
% experiment (sec.~\ref{sub:mutate-interesting}).

>>>>

Putting the rational programmer to work means generating many mutants and turning
those into debugging scenarios. The process must start with a suitable
collection of representative programs (see sec.~\ref{sub:mutate-benchmarks}).
Since existing mutators do not generate useful mutants, the next step is to
develop new mutators (see sec.~\ref{sub:mutate-mutators}) and to validate their
suitability on the benchmarks (see sec.~\ref{sub:mutate-interesting}).

### 2021-05-20 writing/why-rational.tex
move Louisiana footnote so that the intro to the benchmark section makes sense

This paper generalizes Lazarek et. al.'s idea to {\em the rational programmer\/}.
Like~\citet{mill1874essays}'s {\it homo economicus\/}, the rational programmer approximates the
behavior of a software developer who reduces time spent on a task by exploiting the
available information. In the context of gradual typing, the rational programmer
has two pieces of information when an impedance mismatch signals exceptional
behavior: the error message and the state of the program. Hence, the most
rational procedure is to use the former to improve the latter. Specifically,
the rational programmer translates the Wadler-Findler slogan into a debugging
method, searching for the source of the impedance mismatch by adding type
annotations to some of the untyped parts of the program identified in the error.
If the type checker rejects an annotation derived from the context, the
rational programmer has found the source of the problem.
Otherwise, the rational programmer concludes that the just-annotated parts are not the problem
and re-runs the program;
the Wadler-Findler slogan implies the new version of the program must now blame
a different location for the problem, which the rational programmer can check in
the same fashion. 
Measuring this simulated
behavior on a large number of debugging scenarios yields data that
is similar to data collected in a human-facing study.

>>>>

This paper generalizes Lazarek et. al.'s idea to {\em the rational
programmer\/}.  Like~\citet{mill1874essays}'s {\it homo economicus\/}, the
rational programmer approximates the behavior of a software developer who
reduces time spent on a task by exploiting the available information. In the
context of gradual typing, the rational programmer has two pieces of information
when an impedance mismatch signals exceptional behavior: the error message and
the state of the program. Hence, the most rational procedure is to use the
former to improve the latter. Specifically, the rational programmer translates
the Wadler-Findler slogan into a debugging method, searching for the source of
the impedance mismatch by adding type annotations to some of the untyped parts
of the program identified in the error.  If the type checker rejects an
annotation derived from the context, the rational programmer has found the
source of the problem.  Otherwise, the rational programmer concludes that the
just-annotated parts are not the problem and re-runs the program---which must,
by the slogan, blame a different location for the problem. At this point, the
rational programmer can iterate the process.  Measuring this simulated behavior
on a large number of scenarios yields data that is similar to data collected in
a human-facing study.
